Windowed least-squares model reduction for dynamical systems

### 2019-10-06 main.tex
working on EL

\KTC{HERE}
Substituting the approximation~\eqref{eq:stapprox1} into the FOM
ODE~\eqref{eq:FOM} yields
\begin{equation}\label{eq:st_fom}
\stbasisDotArg{}{t} \stgenstate =  \velocity(\stbasisArg{}{t} \stgenstate + \stateIntercept,t) , \qquad \genstate(0) =\mathbb{P}(\stateFOM)(0), \qquad t \in [0,T],
\end{equation}
where 
$\stbasisDot\equiv d\stbasis/d\timeDummy$,
$\mathbb{P} : \RR{N} \otimes \timeSpace \rightarrow \stspaceST$ is an orthogonal projection onto the trial subspace (e.g., $\elltwo$).
Equation~\eqref{eq:st_fom} is an over-determined algebraic system for the $\stdim$ unknowns. We now outline the Space--Time Least-Squares Petrov--Galerkin method~\cite{choi_stlspg}. 
The space--time Galerkin method is noted to additionally be a possibility for model reduction. As the space--time Galerkin method does not associate with any residual minimization 
principle, however, we do not outline it here.
\subsubsection{Space--Time LSPG projection}  
As in LSPG, Space--Time LSPG (ST-LSPG)~\cite{choi_stlspg} minimizes the fully discrete residual, with the key difference being that ST-LSPG minimizes the residual over the entire time domain. For brevity, ST-LSPG is outlined only for linear multistep methods.
%As ST-LSPG acts on the discrete solution, we first note that the discrete FOM solution lives in $\RR{N} \otimes \RR{N_t}$. ST-LSPG then seeks solutions in the trial space,
%$$\stspaceDiscrete \defeq \spaceTrialSpace \otimes \timeTrialSpaceDiscrete \subseteq \RR{N} \otimes \RR{N_t},$$
%where $\timeTrialSpaceDiscrete$ is the discrete temporal trial space. It is noted that this trial space is a specific instance of the trial space~\ref{eq:sttrialspace_def} 
%As before, we define a discretization of the continuous time-domain $[0,T]$ into $N_t$ time instances $t^n = n\Delta t, n=1,2,\ldots,N_t$ with a constant time-step $\Delta t = T/N_t$. We define now the space--time discrete residual for the implicit Euler method as,
We first define the space--time discrete residual for linear multistep methods,
\begin{align*}
\residST_{} & \vcentcolon (\stateyDiscreteArg{N_t}, \ldots ,\stateyDiscreteArg{1};\stateFOMIC) \mapsto \begin{bmatrix}
%\frac{\stateyDiscreteArg{N_t} - \stateyDiscreteArg{N_t-1}  }{\Delta t} -  \velocity(\stateyDiscreteArg{N_t}) \\
\residLMSArg{N_t}(\stateyDiscreteArg{N_t};\stateyDiscreteArg{N_t-1},\ldots,\stateyDiscreteArg{N_t - k^{N_t}}) \\ 
%\frac{\stateyDiscreteArg{N_t-1} - \stateFOMDiscreteArg{N_t-2}  }{\Delta t} -  \velocity(\stateyDiscreteArg{N_t-1}) \\
\vdots \\
%\frac{\stateyDiscreteArg{1} - \stateFOMIC }{\Delta t} -  \velocity(\stateyDiscreteArg{1}) \\
\residLMSArg{1}(\stateyDiscreteArg{1}; \stateFOMIC) 
\end{bmatrix}  \\
& \vcentcolon \RR{\fomdim} \otimes \RR{N_t+1} \rightarrow \RR{\fomdim N_t}. 
\end{align*}
For convenience we define the same residual with the generalized coordinates as the inputs,
\begin{align*}
\genresidST_{\text{}} & \vcentcolon (\stgenstatey ;\stateFOMIC) \mapsto \begin{bmatrix}
%\frac{\stateyDiscreteArg{N_t} - \stateyDiscreteArg{N_t-1}  }{\Delta t} -  \velocity(\stateyDiscreteArg{N_t}) \\
\residLMSArg{n}\bigg( \stbasisArg{}{t^{N_t}}\stgenstatey + \stateIntercept; \stbasisArg{}{t^{N_t-1}}\stgenstatey + \stateIntercept ,\ldots,  \stbasisArg{}{t^{N_t-k^{N_t}}}\stgenstatey + \stateIntercept \bigg) \\ 
\vdots \\
\residLMSArg{n}\bigg( \stbasisArg{}{t^1}\stgenstatey + \stateIntercept; \stateFOMIC \bigg) \\ 
\end{bmatrix}  \\
& \vcentcolon \RR{\stdim} \times \RR{N} \rightarrow \RR{\fomdim N_t}. 
\end{align*}
ST-LSPG is defined as,
\begin{equation*}
\stgenstate = \underset{ \stgenstatey }{\text{arg\,min}}|| \lspgWeightingSTArg{\stgenstatey}  \genresidST_{\text{}}(\stgenstatey;\stateFOMIC) ||_2^2, 
\end{equation*}
where $\lspgWeightingSTArg{\cdot} \in \RR{n_{st} \times N N_t}$, with $\stdim \le n_{st} \le N N_t$, is a space--time weighting matrix.

@@

\section{Derivation of the Euler--Lagrange Equations}\label{appendix:eulerlagrange}
This section details the derivation of the Euler Lagrange equations. To this end, we consider the generic functional of the form,
\begin{equation}\label{eq:el_gen}
\mathcal{J}: (\statey,\stateyDot) \mapsto  \int_a^b \integrand(\statey(t),\stateyDot(t),t)  dt,
where $\integrand: \RR{M} \times \RR{M} \times[a,b] \rightarrow  \RR{}$ (for arbitrary $M$) with $\integrand: (\stateyDiscrete,\stateyDotDiscrete,t) \mapsto \integrand(\stateyDiscrete,\stateyDotDiscrete,t)$. We now introduce the function $\statez: [a,b] \rightarrow \RR{M}$ that is implicitly defined as an extremum of~\eqref{eq:el_gen} subject to the boundary condition $\statez(a) = \statez_a$. 
We introduce an arbitrary function $\variation : [a,b] \rightarrow \RR{M}$ with the boundary condition $\variationArg{a} = \bz$ and define a variation from the extremal by,
\statezBar  &: (\tau;\delta) \mapsto \statez(\tau) + \delta \variation(\tau),\\
&: [a,b] \times \RR{} \rightarrow \RR{M}.
We note that $\statezBar$ obeys the same boundary conditions as $\statez$ since $\variationArg{a} = \bz$.
Evaluating the functional~\eqref{eq:el_gen} at $\statezBar(t,\epsilon)$,

@@

\mathcal{J}(\epsilon) = \int_a^b \integrand(\statezBar(t;\epsilon),\statezDotBar(t;\epsilon) ,t)  dt.

@@

The objective is to now find $\overline{\statez}$ that makes $\mathcal{J}(\epsilon)$ stationary. This can be done by differentiating with respect to $\epsilon$, and setting the result to zero,% Noting that the extermum of Eq.~\ref{eq:el_gen} occurs when $\state = \state^*$, and thus when $\epsilon = 0$, we have,
\frac{d}{d\epsilon} \bigg[ \mathcal{J}\big(\statezBar(t;\epsilon),\statezDotBar(t;\epsilon) \big)\bigg] =  0,
or equivalently,
\int_a^b \frac{d}{d\epsilon} \bigg[ \integrand \big(\statezBar(t;\epsilon),\statezDotBar(t;\epsilon),t \big)\bigg] dt=  0.

@@

$$ \int_a^b \frac{d}{d\epsilon}\bigg[ \integrand \big(\statezBar(t;\epsilon),\statezDotBar(t;\epsilon) ,t\big) \bigg] dt = \int_a^b \bigg[ \frac{\partial \integrand  }{\partial \stateyDiscrete } \big(\statezBar(t;\epsilon),\statezDotBar(t;\epsilon) ,t\big) \frac{\partial \statezBar }{\partial \delta}(t;\epsilon)  + \frac{\partial \integrand}{\partial \stateyDotDiscrete } \big(\statezBar(t;\epsilon),\statezDotBar(t;\epsilon) ,t \big) \frac{\partial \statezDotBar}{\partial \delta} (t;\epsilon)\bigg]dt . $$

@@

$$\frac{\partial \statezBar}{\partial \delta} (\cdot,t)= \variation(t), \qquad \frac{\partial \statezDotBar }{\partial \delta}(\cdot,t)= \variationDot(t),$$

@@

$$\int_a^b \frac{d}{d\epsilon}\bigg[ \integrand \big(\statezBar(t;\epsilon),\statezDotBar(t;\epsilon) ,t \big) \bigg] dt = \int_a^b \bigg[ \frac{\partial \integrand  }{\partial \stateyDiscrete } \big(\statezBar(t;\epsilon),\statezDotBar(t;\epsilon),t \big) \variation(t)  + \frac{\partial \integrand}{\partial \stateyDotDiscrete } \big(\statezBar(t;\epsilon),\statezDotBar(t;\epsilon) ,t \big) \variationDot(t) \bigg]dt. $$

@@

\int_a^b \frac{d}{d\epsilon}\bigg[ \integrand \big(\statezBar(t;\epsilon),\statezDotBar(t;\epsilon) ,t \big) \bigg] dt =  \int_a^b  \bigg[\frac{\partial \integrand  }{\partial \stateyDiscrete } \big(\statezBar(t;\epsilon),\statezDotBar(t;\epsilon) ,t \big) \variation(t) - \frac{d}{dt}\bigg( \frac{\partial \integrand}{\partial \stateyDotDiscrete }  \big(\statezBar(t;\epsilon),\statezDotBar(t;\epsilon) ,t \big) \bigg) \variation(t) \bigg]dt + \\ \frac{\partial \integrand}{\partial \stateyDotDiscrete} \big(\statezBar(b;\epsilon),\statezDotBar(b;\epsilon) ,b \big) \variation(b)  - \frac{\partial \integrand}{\partial \stateyDotDiscrete} \big(\statezBar(a;\epsilon),\statezDotBar(a;\epsilon) ,a\big) \variation(a) .

@@

>>>>

Substituting the approximation~\eqref{eq:stapprox1} into the FOM
ODE~\eqref{eq:FOM} yields
\begin{equation}\label{eq:st_fom}
\stbasisDotArg{}{t} \stgenstate =  \velocity(\stbasisArg{}{t} \stgenstate +
	\stateIntercept,t) , \qquad \genstate(0) =\mathbb{P}(\stateFOM(0)\otimes
	\onesFunction)(0), \qquad t \in [0,T],
\end{equation}
where 
$\stbasisDot\equiv d\stbasis/d\timeDummy$,
$\mathbb{P} : \RR{N} \otimes \timeSpace \rightarrow \stspaceST$ is an orthogonal projection onto the trial subspace (e.g., $\elltwo$).
Equation~\eqref{eq:st_fom} is an over-determined algebraic system for the $\stdim$ unknowns. We now outline the Space--Time Least-Squares Petrov--Galerkin method~\cite{choi_stlspg}. 
The space--time Galerkin method is noted to additionally be a possibility for model reduction. As the space--time Galerkin method does not associate with any residual minimization 
principle, however, we do not outline it here.
\subsubsection{Space--Time LSPG projection}  
As in LSPG, Space--Time LSPG (ST-LSPG)~\cite{choi_stlspg} minimizes the fully discrete residual, with the key difference being that ST-LSPG minimizes the residual over the entire time domain. For brevity, ST-LSPG is outlined only for linear multistep methods.
%As ST-LSPG acts on the discrete solution, we first note that the discrete FOM solution lives in $\RR{N} \otimes \RR{N_t}$. ST-LSPG then seeks solutions in the trial space,
%$$\stspaceDiscrete \defeq \spaceTrialSpace \otimes \timeTrialSpaceDiscrete \subseteq \RR{N} \otimes \RR{N_t},$$
%where $\timeTrialSpaceDiscrete$ is the discrete temporal trial space. It is noted that this trial space is a specific instance of the trial space~\ref{eq:sttrialspace_def} 
%As before, we define a discretization of the continuous time-domain $[0,T]$ into $N_t$ time instances $t^n = n\Delta t, n=1,2,\ldots,N_t$ with a constant time-step $\Delta t = T/N_t$. We define now the space--time discrete residual for the implicit Euler method as,
We first define the space--time discrete residual for linear multistep methods,
\begin{align*}
\residST_{} & \vcentcolon (\stateyDiscreteArg{N_t}, \ldots ,\stateyDiscreteArg{1};\stateFOMIC) \mapsto \begin{bmatrix}
%\frac{\stateyDiscreteArg{N_t} - \stateyDiscreteArg{N_t-1}  }{\Delta t} -  \velocity(\stateyDiscreteArg{N_t}) \\
\residLMSArg{N_t}(\stateyDiscreteArg{N_t};\stateyDiscreteArg{N_t-1},\ldots,\stateyDiscreteArg{N_t - k^{N_t}}) \\ 
%\frac{\stateyDiscreteArg{N_t-1} - \stateFOMDiscreteArg{N_t-2}  }{\Delta t} -  \velocity(\stateyDiscreteArg{N_t-1}) \\
\vdots \\
%\frac{\stateyDiscreteArg{1} - \stateFOMIC }{\Delta t} -  \velocity(\stateyDiscreteArg{1}) \\
\residLMSArg{1}(\stateyDiscreteArg{1}; \stateFOMIC) 
\end{bmatrix}  \\
& \vcentcolon \RR{\fomdim} \otimes \RR{N_t+1} \rightarrow \RR{\fomdim N_t}. 
\end{align*}
For convenience we define the same residual with the generalized coordinates as the inputs,
\begin{align*}
\genresidST_{\text{}} & \vcentcolon (\stgenstatey ;\stateFOMIC) \mapsto \begin{bmatrix}
%\frac{\stateyDiscreteArg{N_t} - \stateyDiscreteArg{N_t-1}  }{\Delta t} -  \velocity(\stateyDiscreteArg{N_t}) \\
\residLMSArg{n}\bigg( \stbasisArg{}{t^{N_t}}\stgenstatey + \stateIntercept; \stbasisArg{}{t^{N_t-1}}\stgenstatey + \stateIntercept ,\ldots,  \stbasisArg{}{t^{N_t-k^{N_t}}}\stgenstatey + \stateIntercept \bigg) \\ 
\vdots \\
\residLMSArg{n}\bigg( \stbasisArg{}{t^1}\stgenstatey + \stateIntercept; \stateFOMIC \bigg) \\ 
\end{bmatrix}  \\
& \vcentcolon \RR{\stdim} \times \RR{N} \rightarrow \RR{\fomdim N_t}. 
\end{align*}
ST-LSPG is defined as,
\begin{equation*}
\stgenstate = \underset{ \stgenstatey }{\text{arg\,min}}|| \lspgWeightingSTArg{\stgenstatey}  \genresidST_{\text{}}(\stgenstatey;\stateFOMIC) ||_2^2, 
\end{equation*}
where $\lspgWeightingSTArg{\cdot} \in \RR{n_{st} \times N N_t}$, with $\stdim \le n_{st} \le N N_t$, is a space--time weighting matrix.

@@

\section{Derivation of the Euler--Lagrange Equations}\label{appendix:eulerlagrange}
This section details the derivation of the Euler Lagrange equations. To this end, we consider the generic functional of the form,
\begin{equation}\label{eq:el_gen}
\mathcal{J}: (\statey,\stateyDot) \mapsto  \int_a^b \integrand(\statey(t),\stateyDot(t),t)  dt,
\end{equation}
where $\integrand: \RR{M} \times \RR{M} \times[a,b] \rightarrow  \RR{}$ (for arbitrary $M$) with $\integrand: (\stateyDiscrete,\stateyDotDiscrete,t) \mapsto \integrand(\stateyDiscrete,\stateyDotDiscrete,t)$. We now introduce the function $\statez: [a,b] \rightarrow \RR{M}$ that is implicitly defined as an extremum of~\eqref{eq:el_gen} subject to the boundary condition $\statez(a) = \statez_a$. 
We introduce an arbitrary function $\variation : [a,b] \rightarrow \RR{M}$ with the boundary condition $\variationArg{a} = \bz$ and define a variation from the extremal by,
\begin{align*}
\statezBar  &: (\tau,\delta) \mapsto \statez(\tau) + \delta \variation(\tau),\\
&: [a,b] \times \RR{} \rightarrow \RR{M}.
\end{align*}
We note that $\statezBar$ obeys the same boundary conditions as $\statez$ since $\variationArg{a} = \bz$.
We define a new function that is equivalent to~\eqref{eq:el_gen} evaluated at $\statezBar(t,\cdot)$,
$$
\mathcal{J}_{\delta}: \delta \mapsto \int_a^b \integrand(\statezBar(t,\delta),\statezDotBar(t,\delta) ,t)  dt.
$$
%\mathcal{J}(\epsilon) = \int_a^b \integrand(\statezBar(t;\epsilon),\statezDotBar(t;\epsilon) ,t)  dt.
The objective is to now find $\overline{\statez}$ that makes $\mathcal{J}_{\delta}$ stationary. This can be done by differentiating with respect to $\delta$, and setting the result to zero,% Noting that the extermum of Eq.~\ref{eq:el_gen} occurs when $\state = \state^*$, and thus when $\epsilon = 0$, we have,

@@

\frac{d}{d\delta} \big( \mathcal{J}_{\delta} \big)=  0.

@@

%or equivalently,
% \int_a^b \frac{d}{d \delta} \big(\integrand(\statezBar(t,\delta),\statezDotBar(t,\delta) ,t)\big)  dt = 0
%\int_a^b \frac{d}{d\epsilon} \bigg[ \integrand \big(\statezBar(t;\epsilon),\statezDotBar(t;\epsilon),t \big)\bigg] dt=  0.

@@

\frac{d}{d\delta} \big( \mathcal{J}_{\delta} \big)(\epsilon)= 
\int_a^b \bigg[ \frac{\partial \integrand  }{\partial \stateyDiscrete } \big(\statezBar(t,\epsilon),\statezDotBar(t,\epsilon) ,t\big) \frac{\partial \statezBar }{\partial \delta}(t,\epsilon)  + \frac{\partial \integrand}{\partial \stateyDotDiscrete } \big(\statezBar(t,\epsilon),\statezDotBar(t,\epsilon) ,t \big) \frac{\partial \statezDotBar}{\partial \delta} (t,\epsilon)\bigg]dt . $$

@@

$$\frac{\partial \statezBar}{\partial \delta} (t,\cdot)= \variation(t), \qquad \frac{\partial \statezDotBar }{\partial \delta}(t,\cdot)= \variationDot(t),$$

@@

%\int_a^b \frac{d}{d\epsilon}\bigg[ \integrand \big(\statezBar(t;\epsilon),\statezDotBar(t;\epsilon) ,t \big) \bigg] dt 
\frac{d}{d\delta} \big( \mathcal{J}_{\delta} \big)(\epsilon)
= \int_a^b \bigg[ \frac{\partial \integrand  }{\partial \stateyDiscrete } \big(\statezBar(t;\epsilon),\statezDotBar(t;\epsilon),t \big) \variation(t)  + \frac{\partial \integrand}{\partial \stateyDotDiscrete } \big(\statezBar(t;\epsilon),\statezDotBar(t;\epsilon) ,t \big) \variationDot(t) \bigg]dt. $$

@@

\frac{d}{d\delta} \big( \mathcal{J}_{\delta} \big)(\epsilon)
%\int_a^b \frac{d}{d\epsilon}\bigg[ \integrand \big(\statezBar(t;\epsilon),\statezDotBar(t;\epsilon) ,t \big) \bigg] dt 
=  \int_a^b  \bigg[\frac{\partial \integrand  }{\partial \stateyDiscrete } \big(\statezBar(t,\epsilon),\statezDotBar(t,\epsilon) ,t \big) \variation(t) - \frac{d}{dt}\bigg( \frac{\partial \integrand}{\partial \stateyDotDiscrete }  \big(\statezBar(t,\epsilon),\statezDotBar(t,\epsilon) ,t \big) \bigg) \variation(t) \bigg]dt + \\ \frac{\partial \integrand}{\partial \stateyDotDiscrete} \big(\statezBar(b,\epsilon),\statezDotBar(b,\epsilon) ,b \big) \variation(b)  - \frac{\partial \integrand}{\partial \stateyDotDiscrete} \big(\statezBar(a,\epsilon),\statezDotBar(a,\epsilon) ,a\big) \variation(a) .

@@

\EP{Here}
